A New Hybrid Machine Learning Method for Stellar Parameter Inference

### 2025-01-19 paper/paper2/paper.tex
Move files

\begin{abstract}
    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
    Here we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
    We first use \texttt{blase}'s \citep{blase} nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines. 
    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 spectral lines.
    For each of these lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra given continuously valued stellar parameters.
    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
    Our products are an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, as well as the spectral reconstruction and inference algorithms themselves.
    This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
\end{abstract}

>>>>

\begin{abstract}
    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
    Here we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
    We first use \texttt{blase}'s \citep{blase} nontraditional semi-empirical approach to treating spectra as sets of tunable spectral lines.
    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 spectral lines.
    For each of these lines, we continuously map stellar parameters to spectral line parameters using regular grid linear interpolation.
    These manifolds are aggregated to create the PHOENIX generator, sporting parallelized reconstruction of spectra given continuously valued stellar parameters.
    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
    Our products are an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, as well as the spectral reconstruction and inference algorithms themselves.
    This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
\end{abstract}
